An Effective Model of HL-60 Differentiation

### 2017-01-23 Paper_v2.tex
Updated results

\section*{Results}
We constructed an effective model of the ATRA-induced HL-60 differentiation which described signaling and gene expression events following the addition of ATRA (Fig. \ref{fig:network}).
The model connectivity was developed from literature and the studies presented here.
We decomposed the ATRA program into three modules;
a signal initiation module that was responsible for sensing and transforming the ATRA signal into activated cRaf-S621 and the ATRA-RXR/RAR (Trigger) signals (Fig. \ref{fig:network}A);
a signal integration module that controlled the expression of upstream transcription factors given cRaf-S621 and activated Trigger (Fig. \ref{fig:network}B); and
a phenotype module which encoded the expression of functional differentiation markers (Fig. \ref{fig:network}C).

>>>>

\section*{Results}
We constructed an effective model of ATRA-induced HL-60 differentiation which described signaling and gene expression events following the addition of ATRA (Fig. \ref{fig:network}).
The model connectivity was developed from literature and the studies presented here (Table ZZ).
We decomposed the ATRA program into three modules;
a signal initiation module that sensed and transformed the ATRA signal into activated cRaf-S621 and the ATRA-RXR/RAR (Trigger) signals (Fig. \ref{fig:network}A);
a signal integration module that controlled the expression of upstream transcription factors given cRaf-S621 and activated Trigger signals (Fig. \ref{fig:network}B); and
a phenotype module which encoded the expression of functional differentiation markers from the ATRA-inducible transcription factors (Fig. \ref{fig:network}C).
Each component of these modules was described by a mRNA and protein balance equation.
Additionally, the signal initiation module also described the abundance of activated species e.g., Trigger and cRaf-S621 whose values were derived from
unactivated Trigger and cRaf protein levels.
Lastly, because the population of HL-60 cells was dividing (at least before ATRA-induced cell cycle arrest), we also considered a dilution term in all balance equations.
The signal initiation module contained nine differential equations, while the signal integration and phenotype modules were collectively encoded by 54 differential equations.
Model parameters were taken literature, or estimated from experimental data taken from literature using heuristic optimization (see materials and methods).
